---
title: Titanic Machine learning from disaster
author: Stijn Hanssen
date: 30 October 2018
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Load packages
Loading the required packages for specific functions.
```{r}
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
pacman::p_load(GLDEX,dplyr,mice,VIM,randomForest)
```

### Set working directory
Choose the path were the data-sets are stored
```{r}
DATA.DIR <- "~/GitHub/Kaggle/TitanicMachineLearningfromDisaster"
setwd(DATA.DIR)
# See which files are in WD
list.files() 
```

### Import data
The training and test set will be imported into R.
```{r}
train <- read.csv("train.csv",stringsAsFactors = F,header = T)
test <- read.csv("test.csv", stringsAsFactors =  F,header = T)
```


### Setting the columns to factors
All these columns have different levels
```{r}
train$Pclass <- as.factor(train$Pclass)
train$Sex <- as.factor(train$Sex)
train$Embarked <- as.factor(train$Embarked)
test$Pclass <- as.factor(test$Pclass)
test$Sex <- as.factor(test$Sex)
test$Embarked <- as.factor(test$Embarked)
```

### Missing values age training
A linear model is used to predict the missing values for the column age. Set up the model then use predict for prediction. Then put the predicted values into the dataset
```{r}
trainage <- train[,c("Age","Pclass","Sex","SibSp","Parch","Fare")]
equation <- "Age ~ ."
formula <- as.formula(equation)
model <- lm(formula = formula,data = trainage)
trainageNA <- which.na(trainage)
trainageNA <- trainage[trainageNA,]
prediction <- predict(model,trainageNA)
prediction
train[is.na(train$Age), "Age"] <- prediction
```

### Missing values embarked training
Assigning the most often occuring character to the two missing values, it is common pratice to replace the missing categorical variables with the mode of the observed onces (however it is questionable).
```{r}
train$Embarked[c(62,830)] <- 'S'
table(train$Embarked)
train$Embarked <- droplevels(train$Embarked)
table(train$Embarked)
```

### Missing value fate test set
For the reason that only one value is missing we used the avarage for this one.
```{r}
which.na(test$Fare)
test$Fare[153] <- mean(test$Fare,na.rm = T)
```

### Missing values age test set
```{r}
testing <- test[,c("Age","Pclass","Sex","SibSp","Parch","Fare")]
equation <- "Age ~ ."
formula <- as.formula(equation)
model <- lm(formula = formula,data = testing)
testingNA <- which.na(testing)
testingNA <- testing[testingNA,]
prediction <- predict(model,testingNA)
prediction
test[is.na(test$Age), "Age"] <- prediction
```

### Survived column training as factor
```{r}
train$Survived <- as.factor(train$Survived)
```

### Prediction model
```{r}
equation <- "Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked"
formula <- as.formula(equation)
model <- randomForest(formula = formula, data = train)
```

### Test model on test data
```{r}
survived <- predict(model, newdata = test)
survived
```

### Kaggle submition document
```{r}
PassengerId <- test$PassengerId
output <- as.data.frame(PassengerId)
output$survived <- survived
tail(output)

write.csv(output, file="kaggle_submission.csv", row.names = F)
```


