---
title: Titanic Machine learning from disaster
author: Stijn Hanssen
date: 30 October 2018
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Load packages
Loading the required packages for specific functions.
```{r}
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
pacman::p_load(GLDEX,dplyr,mice,VIM,randomForest)
```

### Set working directory
Choose the path were the data-sets are stored
```{r}
DATA.DIR <- "~/GitHub/Kaggle/TitanicMachineLearningfromDisaster"
setwd(DATA.DIR)
# See which files are in WD
list.files() 
```

### Import data
The training and test set will be imported into R.
```{r}
train <- read.csv("train.csv",stringsAsFactors = F,header = T)
test <- read.csv("test.csv", stringsAsFactors =  F,header = T)
```

### Inspect training data
Firstly we look at the data structure, to see which variables we have. 
```{r}
str(train)
summary(train)
head(train)
tail(train)
```

### Inspect test data
Firstly we look at the data structure, to see which variables we have. 
```{r}
str(test)
summary(test)
head(test)
tail(test)
```

### Setting the columns to factors
All these columns have different levels
```{r}
train$Pclass <- as.factor(train$Pclass)
train$Sex <- as.factor(train$Sex)
train$Embarked <- as.factor(train$Embarked)
str(train)
test$Pclass <- as.factor(test$Pclass)
test$Sex <- as.factor(test$Sex)
test$Embarked <- as.factor(test$Embarked)
str(test)
```

### Investigate specific columns training set
```{r}
# check at which places there are missing values for age
which.na(train$Age)
# at what place did people entered the ship
table(train$Embarked)
which(train$Embarked == '')
# filter based on paid nothing for the titanic
Fare <- filter(train, Fare == 0.00)
# every person who paided nothing is a man who embarked at S (southampton) it is mentioned that this is the place were the crew boarded
Fare
# cabin column misses alot information
head(train$Cabin, n = 20)
```

### Missing values age training
The data seems to be missing completely at random (MCAR). Usually a safe maximum threshold for missing data is 5 percent of the total for large datasets. Quick fix could be mean-subsitution but this results in decrease of the variance, while the mean will be unchanged.
Mice package imputing missing values with plausible data values (drawn from distribution)
```{r}
# test if more than 5 percent is missing of the age column
pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(train,2,pMiss)
trainage <- train[,c("Age","Pclass","Sex","SibSp","Parch","Fare")]
tempData <- mice(trainage,m=5,maxit=50,meth='pmm',seed=500,printFlag = F)
summary(tempData)
head(tempData$imp$Age)
completedData <- complete(tempData,1)
train$Age <- completedData$Age
which.na(train$Age)
```

### Missing values embarked training
Assigning the most often occuring character to the two missing values, it is common pratice to replace the missing categorical variables with the mode of the observed onces (however it is questionable).
```{r}
train$Embarked[c(62,830)] <- 'S'
table(train$Embarked)
train$Embarked <- droplevels(train$Embarked)
table(train$Embarked)
```

### Missing value fate test set

```{r}
which.na(test$Fare)
test$Fare[153] <- mean(test$Fare,na.rm = T)
test$Fare[153]
```


### Missing values age test set

```{r}
apply(test,2,pMiss)
testing <- test[,c("Age","Pclass","Sex","SibSp","Parch","Fare")]
tempData <- mice(testing,m=5,maxit=50,meth='pmm',seed=500,printFlag = F)
summary(tempData)
tempData$imp$Age
completedData <- complete(tempData,1)
test$Age <- completedData$Age
which.na(test$Age)
```

### Survived column training as factor

```{r}
train$Survived <- as.factor(train$Survived)
```

### Prediction model

```{r}
equation <- "Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked"
formula <- as.formula(equation)
model <- randomForest(formula = formula, data = train, ntree = 500, mtry = 3, nodesize = 0.01 * nrow(train))
```

### Test model on test data
```{r}
survived <- predict(model, newdata = test)
survived
```

### Kaggle submition document

```{r}
PassengerId <- test$PassengerId
output <- as.data.frame(PassengerId)
output$survived <- survived
tail(output)

write.csv(output, file="kaggle_submission.csv", row.names = F)
```


