---
title: BagofWordsMeetsBagsofPopcorn
author: Stijn Hanssen
date: 4 December 2018
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Load packages
Loading the required packages for specific functions.
```{r}
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
pacman::p_load(dplyr,tm,knitr,RWeka,randomForest)
```

### Set working directory
Choose the path were the data-sets are stored
```{r}
DATA.DIR <- "~/GitHub/Kaggle/BagofWordsMeetsBagsofPopcorn"
setwd(DATA.DIR)
# See which files are in WD
list.files() 
```

### Import data
The training and test set will be imported into R.
```{r}
word_train = read.table("labeledTrainData.tsv", header = T)
word_test = read.table("testData.tsv", header = T)
full = bind_rows(word_train,word_test)

dim(word_train)
dim(word_test)
str(full)
```

### Inspect a review
In the text are the line breaks shown as *<br />*. This have to be taken care off. Furthermore, we are not interested in the numberws and punctuations. In addition, we will remove the common stop words, convert the words back to their roots. Lastly, we will remove all the whitespaces.
```{r}
as.character(full$review[[1]])
```


