---
title: "House Prices"
author: "Stijn Hanssen"
date: "1 November 2018"
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Load packages
Loading the required packages for specific functions. If the packaged aren't present, they will be automatically be downloaded.
ggplot for data visualization, stringr for extracting string paterns, matrix for matrix transformations, glmnet for rigde lass and elastinet, xgboost for gbm, metrics for rmse, caret for one hot encoding, e1071 for skewness, corrplot for correlation plot.
```{r}
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
pacman::p_load("stringr","GLDEX","RSEIS","DescTools","caret","ggplot2","Matrix","glmnet","xgboost","randomForest","Metrics","dplyr","scales","e1071","corrplot","GGally")
```

### Set working directory
Choose the path were the data-sets are stored
```{r}
DATA.DIR <- "~/GitHub/Kaggle/HousePricesAdvancedRegressionTechniques"
setwd(DATA.DIR)
# See which files are in WD
list.files() 
```

### Import data
combine the datasets (remove the id from both datasets and remove the saleprice from the train set)
```{r}
train <- read.csv("train.csv",stringsAsFactors = F,header = T)
test <- read.csv("test.csv", stringsAsFactors =  F,header = T)
df.combined <- rbind(within(train, rm('Id','SalePrice')), within(test, rm('Id')))
dim(df.combined)
```

## Check NA
```{r}
na.cols <- which(colSums(is.na(df.combined)) > 0)
sort(colSums(sapply(df.combined[na.cols], is.na)), decreasing = TRUE)
```

## Plotting categorical data
```{r}
plot.categoric <- function(cols, df){
  for (col in cols) {
    order.cols <- names(sort(table(df.combined[,col]), decreasing = TRUE))
  
    num.plot <- qplot(df[,col]) +
      geom_bar(fill = 'cornflowerblue') +
      geom_text(aes(label = ..count..), stat='count', vjust=-0.5) +
      theme_minimal() +
      scale_y_continuous(limits = c(0,max(table(df[,col]))*1.1)) +
      scale_x_discrete(limits = order.cols) +
      xlab(col) +
      theme(axis.text.x = element_text(angle = 30, size=12))
  
    print(num.plot)
  }
}
```

## Pool quality
10 house do have a pool. Check if houses with NA for Pool quality indeed have a poolArea set as 0. 3 houses have a pool area but not a pool quality. we assign these by mean of the known pools. the rest of the pools will be assigned 'None'.
```{r}
plot.categoric('PoolQC', df.combined)
df.combined[(df.combined$PoolArea > 0) & is.na(df.combined$PoolQC),c('PoolQC','PoolArea')]
df.combined[,c('PoolQC','PoolArea')] %>%
  group_by(PoolQC) %>%
  summarise(mean = mean(PoolArea), counts = n()) 
df.combined[2421,'PoolQC'] = 'Ex'
df.combined[2504,'PoolQC'] = 'Ex'
df.combined[2600,'PoolQC'] = 'Fa'
df.combined$PoolQC[is.na(df.combined$PoolQC)] = 'None'
```

## Garage
GarageType: Garage location
GarageYrBlt: Year garage was built
GarageFinish: Interior finish of the garage
GarageCars: Size of garage in car capacity
GarageArea: Size of garage in square feet
GarageQual: Garage quality
GarageCond: Garage condition

How many houses were built the same year their garage was built. Most of this datasets house were built at the same year so we can asume that the NA's did the same.
```{r}
length(which(df.combined$GarageYrBlt == df.combined$YearBuilt))
idx <- which(is.na(df.combined$GarageYrBlt))
df.combined[idx, 'GarageYrBlt'] <- df.combined[idx, 'YearBuilt']
```

check if garageArea and Cars are 0 and if the rest of the garage features have Na on these rows we can set them to 'None', assuming that they dont have a garage. For the house with GarageArea = 360 and GarageCars = 1, but NA's in the other columns, we can use the most frequent values for each columns from houses with a similar area and car count.
Now we can fill in any missing numeric values with 0 and categoric with 'None' since these houses recorded having 0 area and 0 cars in their garage.
```{r}
garage.cols <- c('GarageArea', 'GarageCars', 'GarageQual', 'GarageFinish', 'GarageCond', 'GarageType')
df.combined[is.na(df.combined$GarageCond),garage.cols]

idx <- which(((df.combined$GarageArea < 370) & (df.combined$GarageArea > 350)) & (df.combined$GarageCars == 1))
names(sapply(df.combined[idx, garage.cols], function(x) sort(table(x), decreasing=TRUE)[1]))
df.combined[2127,'GarageQual'] = 'TA'
df.combined[2127, 'GarageFinish'] = 'Unf'
df.combined[2127, 'GarageCond'] = 'TA'

for (col in garage.cols){
  if (sapply(df.combined[col], is.numeric) == TRUE){
    df.combined[sapply(df.combined[col], is.na), col] = 0
  }
  else{
    df.combined[sapply(df.combined[col], is.na), col] = 'None'
  }
}
```

## Kitchen and Eletrical
KitchenQual: Kitchen quality
Electrical: Electrical system
```{r}
plot.categoric('KitchenQual', df.combined)
df.combined$KitchenQual[is.na(df.combined$KitchenQual)] = 'TA'

plot.categoric('Electrical', df.combined)
df.combined$Electrical[is.na(df.combined$Electrical)] = 'SBrkr'
```

## Basement
BsmtQual: Height of the basement
BsmtCond: General condition of the basement
BsmtExposure: Walkout or garden level basement walls
BsmtFinType1: Quality of basement finished area
BsmtFinSF1: Type 1 finished square feet
BsmtFinType2: Quality of second finished area (if present)
BsmtFinSF2: Type 2 finished square feet
BsmtUnfSF: Unfinished square feet of basement area
TotalBsmtSF: Total square feet of basement area
BsmtFullBath: Basement full bathrooms
BsmtHalfBath: Basement half bathrooms

Subset the basement. Almost all of the missing values for each categoric basement feature comes from houses with 0 on each features corresponding to area. We can fill in these values with 'None' since these houses certainly don't have basements. Rows 949, 1488 and 2349 are the only missing values from BsmtExposure, we can fill this with No as that is the most frequent value and these houses most likely don't have any exposure for their basements. 
```{r}
bsmt.cols <- names(df.combined)[sapply(names(df.combined), function(x) str_detect(x, 'Bsmt'))]

df.combined[is.na(df.combined$BsmtExposure),bsmt.cols]
plot.categoric('BsmtExposure', df.combined)
df.combined[c(949, 1488, 2349), 'BsmtExposure'] = 'No'

for (col in bsmt.cols){
  if (sapply(df.combined[col], is.numeric) == TRUE){
    df.combined[sapply(df.combined[col], is.na),col] = 0
  }
  else{
    df.combined[sapply(df.combined[col],is.na),col] = 'None'
  }
}
```

## Exterior
Exterior1st: Exterior covering on house
Exterior2nd: Exterior covering on house (if more than one material)
There is only 1 missing value for Exterior1st and Exterior2nd coming from the same hosue. We can fill this with 'Other' since the NA is likely due to having an exterior cover that is not listed.
```{r}
plot.categoric(c('Exterior1st', 'Exterior2nd'), df.combined)
idx <- which(is.na(df.combined$Exterior1st) | is.na(df.combined$Exterior2nd))
df.combined[idx,c('Exterior1st', 'Exterior2nd')]
df.combined$Exterior1st[is.na(df.combined$Exterior1st)] = 'Other'
df.combined$Exterior2nd[is.na(df.combined$Exterior2nd)] = 'Other'
```

## Type sale, home functionality rating, utilities
SaleType: Type of sale
Functional: Home functionality rating
Utilities: Type of utilities available

SaleType we can see what the SaleCondition of the house was and use a contingency table to see which SaleType and SaleCondition overlap together the most. Most houses with a SaleCondition of 'Normal' almost all have a SaleType of 'WD'. We'll replace the missing value accordingly.
Utilities only has 1 value for NoSeWa and the rest AllPub. We can drop this feature from our dataset as the house with 'NoSeWa' is from our training set and will have won't help with any predictive modelling
```{r}
plot.categoric('SaleType', df.combined)
df.combined[is.na(df.combined$SaleType),c('SaleCondition')]
table(df.combined$SaleCondition, df.combined$SaleType)
df.combined$SaleType[is.na(df.combined$SaleType)] = 'WD'

plot.categoric('Functional', df.combined)
df.combined$Functional[is.na(df.combined$Functional)] = 'Typ'

plot.categoric('Utilities', df.combined)
which(df.combined$Utilities == 'NoSeWa')
col.drops <- c('Utilities')
df.combined <- df.combined[,!names(df.combined) %in% c('Utilities')]
```

## MSZoning and MSSUBclass
We can see what the subclass is for the houses with missing values for Zoning. For Subclasses with 20 'RL' has the largest frequency, however, for Subclasses with 30 and 70 'RM' has the most frequency. We will fill the missing values accordingly.
```{r}
df.combined[is.na(df.combined$MSZoning),c('MSZoning','MSSubClass')]
plot.categoric('MSZoning', df.combined)
table(df.combined$MSZoning, df.combined$MSSubClass)
df.combined$MSZoning[c(2217, 2905)] = 'RL'
df.combined$MSZoning[c(1916, 2251)] = 'RM'
```

## Masonry
For the house with a MasVnrArea of 198 but NA for MasVnrType we can record the median areas for each type and see which type is closest to 198
```{r}
df.combined[(is.na(df.combined$MasVnrType)) | (is.na(df.combined$MasVnrArea)), c('MasVnrType', 'MasVnrArea')]
na.omit(df.combined[,c('MasVnrType','MasVnrArea')]) %>%
  group_by(na.omit(MasVnrType)) %>%
  summarise(MedianArea = median(MasVnrArea,na.rm = TRUE), counts = n()) %>%
  arrange(MedianArea)
plot.categoric('MasVnrType', df.combined)
df.combined[2611, 'MasVnrType'] = 'BrkFace'
df.combined$MasVnrType[is.na(df.combined$MasVnrType)] = 'None'
df.combined$MasVnrArea[is.na(df.combined$MasVnrArea)] = 0
```

## Lotfrontage
There are 486 missing values for LotFrontage, which is quite a lot of values to fill. "LotFrontage: Linear feet of street connected to property." The area of each street connected to the house property is most likely going to have a similar area to other houses in its neighborhood. We can group by each neighborhood and take the median of each LotFrontage and fill the missing values of each LotFrontage based on what neighborhood the house comes from.
```{r}
df.combined['Nbrh.factor'] <- factor(df.combined$Neighborhood, levels = unique(df.combined$Neighborhood))

lot.by.nbrh <- df.combined[,c('Neighborhood','LotFrontage')] %>%
  group_by(Neighborhood) %>%
  summarise(median = median(LotFrontage, na.rm = TRUE))
lot.by.nbrh

idx = which(is.na(df.combined$LotFrontage))

for (i in idx){
  lot.median <- lot.by.nbrh[lot.by.nbrh == df.combined$Neighborhood[i],'median']
  df.combined[i,'LotFrontage'] <- lot.median[[1]]
}
```

## Fence Quality and MiscFeature

```{r}
plot.categoric('Fence', df.combined)
df.combined$Fence[is.na(df.combined$Fence)] = 'None'

table(df.combined$MiscFeature)
df.combined$MiscFeature[is.na(df.combined$MiscFeature)] = 'None'
```

## Fireplaces
Fireplaces: Number of fireplaces
FireplaceQu: Fireplace quality

All the houses that have missing values did not record having any fireplaces. We can replace the NA's with 'None' since these houses don't have any fireplaces at all.
```{r}
plot.categoric('FireplaceQu', df.combined)
which((df.combined$Fireplaces > 0) & (is.na(df.combined$FireplaceQu)))
df.combined$FireplaceQu[is.na(df.combined$FireplaceQu)] = 'None'
```

## Alley

```{r}
plot.categoric('Alley', df.combined)
df.combined$Alley[is.na(df.combined$Alley)] = 'None'
```

## Custom numeric features

Split our data into a numeric set and a categoric set. Transform any categoric into a binary feature using one-hot encoding.
```{r}
num_features <- names(which(sapply(df.combined, is.numeric)))
cat_features <- names(which(sapply(df.combined, is.character)))

df.numeric <- df.combined[num_features]
```

transform any of the ordinal variables (variables that can be scaled) into numeric values. 
Determining which order the categories follow and assigning the values an order from 1,2,..,n.
```{r}
group.df <- df.combined[1:1460,]
group.df$SalePrice <- train$SalePrice

# function that groups a column by its features and returns the mdedian saleprice for each unique feature. 
group.prices <- function(col) {
  group.table <- group.df[,c(col, 'SalePrice', 'OverallQual')] %>%
    group_by_(col) %>%
    summarise(mean.Quality = round(mean(OverallQual),2),
      mean.Price = mean(SalePrice), n = n()) %>%
    arrange(mean.Quality)
    
  print(qplot(x=reorder(group.table[[col]], -group.table[['mean.Price']]), y=group.table[['mean.Price']]) +
    geom_bar(stat='identity', fill='cornflowerblue') +
    theme_minimal() +
    scale_y_continuous(labels = dollar) +
    labs(x=col, y='Mean SalePrice') +
    theme(axis.text.x = element_text(angle = 45)))
  
  return(data.frame(group.table))
}

## functional to compute the mean overall quality for each quality
quality.mean <- function(col) {
  group.table <- df.combined[,c(col, 'OverallQual')] %>%
    group_by_(col) %>%
    summarise(mean.qual = mean(OverallQual)) %>%
    arrange(mean.qual)
  
  return(data.frame(group.table))
}


# function that maps a categoric value to its corresponding numeric value and returns that column to the data frame
map.fcn <- function(cols, map.list, df){
  for (col in cols){
    df[col] <- as.numeric(map.list[df.combined[,col]])
  }
  return(df)
}

```

Any of the columns with the suffix 'Qual' or 'Cond' denote the quality or condition of that specific feature. Each of these columns have the potential values: TA, Fa, Gd, None, Ex, Po. We'll compute the mean house prices for these unique values to get a better sense of what their abbreviations mean.
```{r}
qual.cols <- c('ExterQual', 'ExterCond', 'GarageQual', 'GarageCond', 'FireplaceQu', 'KitchenQual', 'HeatingQC', 'BsmtQual')

group.prices('FireplaceQu')
group.prices('BsmtQual')
group.prices('KitchenQual')

```
From seeing the mean saleprices from a few of the quality and condition features we can infer that the abbreviations mean poor, fair, typical/average, good and excelent. We'll map numeric values from 0-5 to their corresponding categoric values (including 0 for None) and combine that to our dataframe.
```{r}
qual.list <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)

df.numeric <- map.fcn(qual.cols, qual.list, df.numeric)
```

```{r}
group.prices('BsmtExposure')
bsmt.list <- c('None' = 0, 'No' = 1, 'Mn' = 2, 'Av' = 3, 'Gd' = 4)

df.numeric = map.fcn(c('BsmtExposure'), bsmt.list, df.numeric)
group.prices('BsmtFinType1')

```

```{r}
# visualization for BsmtFinTyp2 instead of another table
df.combined[,c('BsmtFinType1', 'BsmtFinSF1')] %>%
  group_by(BsmtFinType1) %>%
  summarise(medianArea = median(BsmtFinSF1), counts = n()) %>%
  arrange(medianArea) %>%
  ggplot(aes(x=reorder(BsmtFinType1,-medianArea), y=medianArea)) +
  geom_bar(stat = 'identity', fill='cornflowerblue') +
  labs(x='BsmtFinType2', y='Median of BsmtFinSF2') +
  geom_text(aes(label = sort(medianArea)), vjust = -0.5) +
  scale_y_continuous(limits = c(0,850)) +
  theme_minimal()
```

```{r}
bsmt.fin.list <- c('None' = 0, 'Unf' = 1, 'LwQ' = 2,'Rec'= 3, 'BLQ' = 4, 'ALQ' = 5, 'GLQ' = 6)
df.numeric <- map.fcn(c('BsmtFinType1','BsmtFinType2'), bsmt.fin.list, df.numeric)
```

This feature doesn't really tell us much and functionality is very vague to tie which other features have a correlation with it. We can compute the mean sale prices for each functional category and assign numeric values accordingly.
```{r}
group.prices('Functional')
functional.list <- c('None' = 0, 'Sal' = 1, 'Sev' = 2, 'Maj2' = 3, 'Maj1' = 4, 'Mod' = 5, 'Min2' = 6, 'Min1' = 7, 'Typ'= 8)

df.numeric['Functional'] <- as.numeric(functional.list[df.combined$Functional])
```

```{r}
group.prices('GarageFinish')
garage.fin.list <- c('None' = 0,'Unf' = 1, 'RFn' = 1, 'Fin' = 2)

df.numeric['GarageFinish'] <- as.numeric(garage.fin.list[df.combined$GarageFinish])


group.prices('Fence')
fence.list <- c('None' = 0, 'MnWw' = 1, 'GdWo' = 1, 'MnPrv' = 2, 'GdPrv' = 4)

df.numeric['Fence'] <- as.numeric(fence.list[df.combined$Fence])

MSdwelling.list <- c('20' = 1, '30'= 0, '40' = 0, '45' = 0,'50' = 0, '60' = 1, '70' = 0, '75' = 0, '80' = 0, '85' = 0, '90' = 0, '120' = 1, '150' = 0, '160' = 0, '180' = 0, '190' = 0)

df.numeric['NewerDwelling'] <- as.numeric(MSdwelling.list[as.character(df.combined$MSSubClass)])
```

We're particularly interested in variables that show strong relationship with SalePrice so we will focus primarily on features that have a coefficient > .5 or < -.5.
```{r}
# need the SalePrice column
corr.df <- cbind(df.numeric[1:1460,], train['SalePrice'])

# only using the first 1460 rows - training data
correlations <- cor(corr.df)
# only want the columns that show strong correlations with SalePrice
corr.SalePrice <- as.matrix(sort(correlations[,'SalePrice'], decreasing = TRUE))

corr.idx <- names(which(apply(corr.SalePrice, 1, function(x) (x > 0.5 | x < -0.5))))

corrplot(as.matrix(correlations[corr.idx,corr.idx]), type = 'upper', method='color', addCoef.col = 'black', tl.cex = .7,cl.cex = .7, number.cex=.7)
```

We can print a matrix of scatter plots to see what these relationships look like under the hood to get a better sense of whats going on.
```{r}
lm.plt <- function(data, mapping, ...){
   plt <- ggplot(data = data, mapping = mapping) + 
    geom_point(shape = 20, alpha = 0.7, color = 'darkseagreen') +
    geom_smooth(method=loess, fill="red", color="red") +
    geom_smooth(method=lm, fill="blue", color="blue") +
    theme_minimal()
  return(plt)
}

ggpairs(corr.df, corr.idx[1:6], lower = list(continuous = lm.plt))
```
The blue lines in the scatter plots represent a simple linear regression fit while the red lines represent a local polynomial fit. We can see both OverallQual and GrLivArea and TotalBsmtSF follow a linear model, but have some outliers we may want to look into. For instance, there are multiple houses with an overall quality of 10, but have suspisciously low prices. We can see similar behavior in GrLivArea and TotalBsmtSF. GarageCars and GarageArea both follow more of a quadratic fit. It seems that having a 4 car garage does not result in a higher house price and same with an extremely large area.
```{r}
ggpairs(corr.df, corr.idx[c(1,7:11)], lower = list(continuous = lm.plt))
```
More of the same with the remaining features with 1stFlrSF, FullBath, TotRmsAbvGrd following linear model while YearBuilt and YearRemodAdd both having non-linear/quadratic models. Taking a closer look to YearBuilt and YearRemodAdd we can see that the most expensive homes are the most recently built and remodelled.
```{r}
plot.categoric('LotShape', df.combined)
```
LotShape has 3 values for having an irregular shape and only 1 for regular. We can create a binary column that returns 1 for houses with a regular lot shape and 0 for houses with any of the 3 irregular lot shapes. Using this method of turning a categoric feature into a binary column will ultimately help our data train better through boosted models without using numeric placeholders on nominal data.
```{r}
df.numeric['RegularLotShape'] <- (df.combined$LotShape == 'Reg') * 1
```
We'll use this exact same method for:

LandContour: Flatness of the property
LandSlope: Slope of property
Electrical: Electrical system
GarageType: Garage location
PavedDrive: Paved driveway
MiscFeature: Miscellaneous feature not covered in other categories
```{r}
plot.categoric('LandContour', df.combined)
df.numeric['LandLeveled'] <- (df.combined$LandContour == 'Lvl') * 1

plot.categoric('LandSlope', df.combined)
df.numeric['LandSlopeGentle'] <- (df.combined$LandSlope == 'Gtl') * 1

plot.categoric('Electrical', df.combined)
df.numeric['ElectricalSB'] <- (df.combined$Electrical == 'SBrkr') * 1

plot.categoric('GarageType', df.combined)
df.numeric['GarageDetchd'] <- (df.combined$GarageType == 'Detchd') * 1

plot.categoric('PavedDrive', df.combined)
df.numeric['HasPavedDrive'] <- (df.combined$PavedDrive == 'Y') * 1

df.numeric['HasWoodDeck'] <- (df.combined$WoodDeckSF > 0) * 1

df.numeric['Has2ndFlr'] <- (df.combined$X2ndFlrSF > 0) * 1

df.numeric['HasMasVnr'] <- (df.combined$MasVnrArea > 0) * 1

```

For MiscFeature the only feature with a significant amount of houses having it is Shed. We can one-hot encode houses that have Sheds vs those who do not
```{r}
plot.categoric('MiscFeature', df.combined)
df.numeric['HasShed'] <- (df.combined$MiscFeature == 'Shed') * 1
```

Many of the houses recorded the same year for YearBuilt and YearRemodAdd. We can create a new column that records that a house was remodelled if the year it was built is different than the remodel year. 
```{r}
df.numeric['Remodeled'] <- (df.combined$YearBuilt != df.combined$YearRemodAdd) * 1
```

We can also create a column that seperates which houses have been recently remodelled vs those who are not. Houses that have been remodelled after the year they were sold will fall into this category.
```{r}
df.numeric['RecentRemodel'] <- (df.combined$YearRemodAdd >= df.combined$YrSold) * 1
```

There can be potential value to homes who were sold the same year they were built as this could be an indicator that these houses were hot in the market.
```{r}
df.numeric['NewHouse'] <- (df.combined$YearBuilt == df.combined$YrSold) * 1
```

What about the houses with area based features equal to 0? Houses with 0 square footage for a columnshows that the house does not have that feature at all. We add a one-hot encoded column for returning 1 for any house with an area greater than 0 since this means that the house does have this feature and 0 for those who do not!
```{r}
#cols.binary <- c('X2ndFlrSF', 'MasVnrArea', 'WoodDeckSF')
cols.binary <- c('X2ndFlrSF', 'MasVnrArea', 'WoodDeckSF', 'OpenPorchSF', 'EnclosedPorch', 'X3SsnPorch', 'ScreenPorch')

for (col in cols.binary){
  df.numeric[str_c('Has',col)] <- (df.combined[,col] != 0) * 1
}
```

We know how important the year a house was built and sold but what about what the specific month it was sold? How do houses sold during summer compare to the other seasons?
```{r}
ggplot(df.combined, aes(x=MoSold)) +
  geom_bar(fill = 'cornflowerblue') +
  geom_text(aes(label=..count..), stat='count', vjust = -.5) +
  theme_minimal() +
  scale_x_continuous(breaks = 1:12)
```

The largest proportion of houses sold is during the summer months: May, June, July. Let's add a column that seperates the the summer houses from the rest.
```{r}
df.numeric['HighSeason'] <- (df.combined$MoSold %in% c(5,6,7)) * 1
```

What about which Neighborhoods are more expensive than others?
```{r}
train[,c('Neighborhood','SalePrice')] %>%
  group_by(Neighborhood) %>%
  summarise(median.price = median(SalePrice, na.rm = TRUE)) %>%
  arrange(median.price) %>%
  mutate(nhbr.sorted = factor(Neighborhood, levels=Neighborhood)) %>%
  ggplot(aes(x=nhbr.sorted, y=median.price)) +
  geom_point() +
  geom_text(aes(label = median.price, angle = 45), vjust = 2) +
  theme_minimal() +
  labs(x='Neighborhood', y='Median price') +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45))
```

StoneBr, NoRidge, NridgHt have a large gap between them versus the rest of the median prices from any of the other neighborhods. It would be wise of us to check if this is from outliers or if these houses are much pricier as a whole.
```{r}
other.nbrh <- unique(df.combined$Neighborhood)[!unique(df.combined$Neighborhood) %in% c('StoneBr', 'NoRidge','NridgHt')]

ggplot(train, aes(x=SalePrice, y=GrLivArea, colour=Neighborhood)) +
  geom_point(shape=16, alpha=.8, size=4) +
  scale_color_manual(limits = c(other.nbrh, 'StoneBr', 'NoRidge', 'NridgHt'), values = c(rep('black', length(other.nbrh)), 'indianred',
                                    'cornflowerblue', 'darkseagreen')) +
  theme_minimal() +
  scale_x_continuous(label=dollar)
```

3 houses from StoneBr, NoRidge and NridgHt with house prices over $500,000 which no other Neighborhood is in the same range. What pops out even more in this plot is the 2 houses with an enormous GrLivArea of over 4500 square ft yet very low sale prices. We certianly have an opportunity to explore these outliers in depth, we'll take a look at this later.
```{r}
nbrh.rich <- c('Crawfor', 'Somerst, Timber', 'StoneBr', 'NoRidge', 'NridgeHt')
df.numeric['NbrhRich'] <- (df.combined$Neighborhood %in% nbrh.rich) *1
```

How about a numeric mapping to the neighborhoods who have higher quality homes and run for larger sale prices?
```{r}
group.prices('Neighborhood')
nbrh.map <- c('MeadowV' = 0, 'IDOTRR' = 1, 'Sawyer' = 1, 'BrDale' = 1, 'OldTown' = 1, 'Edwards' = 1, 
             'BrkSide' = 1, 'Blueste' = 1, 'SWISU' = 2, 'NAmes' = 2, 'NPkVill' = 2, 'Mitchel' = 2,
             'SawyerW' = 2, 'Gilbert' = 2, 'NWAmes' = 2, 'Blmngtn' = 2, 'CollgCr' = 2, 'ClearCr' = 3, 
             'Crawfor' = 3, 'Veenker' = 3, 'Somerst' = 3, 'Timber' = 3, 'StoneBr' = 4, 'NoRidge' = 4, 
             'NridgHt' = 4)

df.numeric['NeighborhoodBin'] <- as.numeric(nbrh.map[df.combined$Neighborhood])
```

Condition of sale
```{r}
group.prices('SaleCondition')
df.numeric['PartialPlan'] <- (df.combined$SaleCondition == 'Partial') * 1
```

```{r}
group.prices('HeatingQC')
heating.list <- c('Po' = 0, 'Fa' = 1, 'TA' = 2, 'Gd' = 3, 'Ex' = 4)

df.numeric['HeatingScale'] <- as.numeric(heating.list[df.combined$HeatingQC])
```

```{r}
area.cols <- c('LotFrontage', 'LotArea', 'MasVnrArea', 'BsmtFinSF1', 'BsmtFinSF2', 'BsmtUnfSF',
             'TotalBsmtSF', 'X1stFlrSF', 'X2ndFlrSF', 'GrLivArea', 'GarageArea', 'WoodDeckSF', 
             'OpenPorchSF', 'EnclosedPorch', 'X3SsnPorch', 'ScreenPorch', 'LowQualFinSF', 'PoolArea')

df.numeric['TotalArea'] <- as.numeric(rowSums(df.combined[,area.cols]))

df.numeric['AreaInside'] <- as.numeric(df.combined$X1stFlrSF + df.combined$X2ndFlrSF)
```

We've seen how strong of an effect the year of a house built has on the house price, therefore, as this dataset collects houses up until 2010 we can determine how old a house is and how long ago the house was sold:
```{r}
df.numeric['Age'] <- as.numeric(2010 - df.combined$YearBuilt)

df.numeric['TimeSinceSold'] <- as.numeric(2010 - df.combined$YrSold)

# how many years since the house was remodelled and sold 
df.numeric['YearSinceRemodel'] <- as.numeric(df.combined$YrSold - df.combined$YearRemodAdd)
```

Correlation plot with OverallQual
```{r}
corr.OverallQual <- as.matrix(sort(correlations[,'OverallQual'], decreasing = TRUE))

corr.idx <- names(which(apply(corr.OverallQual, 1, function(x) (x > 0.5 | x < -0.5))))

corrplot(as.matrix(correlations[corr.idx, corr.idx]), type = 'upper',
         method = 'color', addCoef.col = 'black', tl.cex =.7, cl.cex = .7,
         number.cex = .7)
```

## Outliers
Earlier we saw some suspicious houses with abnormally large GrLivArea's, 2 of which had very low SalePrices. These outliers may disrupt our ability to accurately predict. Lets take a closer look at these values and see if we may want to remove them
```{r}
train.test.df <- rbind(dplyr::select(train,-SalePrice), test)
train.test.df$type <- c(rep('train',1460),rep('test',1459))

ggplot(train, aes(x=GrLivArea)) +
  geom_histogram(fill='lightblue',color='white') +
  theme_minimal()
```

```{r}
outlier_values <- boxplot.stats(train$GrLivArea)$out  # outlier values.
boxplot(train$GrLivArea, main="GrLivArea", boxwex=0.1)
mtext(paste("Outliers: ", paste(outlier_values[outlier_values>4000], collapse=", ")), cex=0.6)
```

```{r}
ggplot(train.test.df, aes(x=type, y=GrLivArea, fill=type)) +
  geom_boxplot() +
  theme_minimal() +
  scale_fill_manual(breaks = c("test", "train"), values = c("indianred", "lightblue"))
```

For the training data we can see 4 houses whose GrLivArea is greater than 4000 yet there is one in the testing set. While it is always a great option to remove outliers from our data as they are usually telling us something more than meets the eye about what is going on with the data, these houses in the trainng set are obnoxiously large and ultimately do not add much value and are causing heavy skewness in both the SalePrice and GrLivArea and particularly the 2 values that have above a 4000 GrLivArea but low SalePrice are putting a constraint on the correlation between the 2 variables. Lets throw these houses out of training set begin preprocessing our data now that there are no heavy outliers.
```{r}
idx.outliers <- which(train$GrLivArea > 4000)
df.numeric <- df.numeric[!1:nrow(df.numeric) %in% idx.outliers,]
df.combined <- df.combined[!1:nrow(df.combined) %in% idx.outliers,]
dim(df.numeric)
```

## Preprocessing
