---
title: "House Prices"
author: "Stijn Hanssen"
date: "15 November 2018"
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Load packages
Loading the required packages for specific functions. If the packaged aren't present, they will be automatically be downloaded.
```{r}
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
pacman::p_load("stringr","GLDEX","RSEIS","DescTools","caret","plyr","Metrics","knitr","ggplot2","plyr","dplyr","corrplot","caret","gridExtra","scales","Rmisc","ggrepel","randomForest","psych","xgboost")
```

### Set working directory
Choose the path were the data-sets are stored.
```{r}
DATA.DIR <- "~/GitHub/Kaggle/HousePricesAdvancedRegressionTechniques"
setwd(DATA.DIR)
# See which files are in WD
list.files() 
```

### Import data
The training and test set will be imported. stringsAsFactors will be set to False in order to prevent that the character columns will be set to factor variables.
```{r}
train <- read.csv("train.csv",stringsAsFactors = F)
test <- read.csv("test.csv", stringsAsFactors =  F)
```

### Check datasets
The structure shows that the variables are integers or characters. 
```{r}
str(train)
```

