---
title: "House Prices"
author: "Stijn Hanssen"
date: "15 November 2018"
output: html_document
---

### Clear working space
First step is to make sure that the working space is empty.
```{r}
# clear all variables
rm(list=ls()) 
```

### Load packages
Loading the required packages for specific functions. If the packaged aren't present, they will be automatically be downloaded.
```{r}
if (!require("pacman")) suppressPackageStartupMessages(install.packages("pacman"))
pacman::p_load("stringr","GLDEX","RSEIS","DescTools","caret","plyr","Metrics","knitr","ggplot2","plyr","dplyr","corrplot","caret","gridExtra","scales","Rmisc","ggrepel","randomForest","psych","xgboost")
```

### Set working directory
Choose the path were the data-sets are stored.
```{r}
DATA.DIR <- "~/GitHub/Kaggle/HousePricesAdvancedRegressionTechniques"
setwd(DATA.DIR)
# See which files are in WD
list.files() 
```

### Import data
The training and test set will be imported. stringsAsFactors will be set to False in order to prevent that the character columns will be set to factor variables.
```{r}
train <- read.csv("train.csv",stringsAsFactors = F)
test <- read.csv("test.csv", stringsAsFactors =  F)
```

### Check and edit the datasets
The structure shows that the variables are integers or characters. Furthermore, the iD variables can be removed from both datasets. But for the prediction of the test set we will need the Id variable thus those will be saved. The test set does not contain Saleprices, thus we will create it and give it "Na" values. Now that the datasets have equal rows and columns they will be combined.
```{r}
str(train)

train$Id = NULL
testID = test$Id
test$Id = NULL

test$SalePrice = NA

df.combined = rbind(train,test)
dim(df.combined)
```

